---
title: "The United Nations voting dataset"
author: "David Aponte"
date: "July 30, 2018"
output: html_document
---


Exploration into UN voting. This markdown file will go through cleaning, EDA, and modeling.

```{r}

library(tidyverse)

votes

# Filter for votes that are "yes", "abstain", or "no"
votes %>%
  filter(vote <= 3)

# Load the countrycode package
library(countrycode)

# Convert country code 100
countrycode(100, "cown", "country.name")

# Add a country column within the mutate: votes_processed
votes_processed <- votes %>%
  filter(vote <= 3) %>%
  mutate(year = session + 1945, country = countrycode(ccode, "cown", "country.name"))

by_year <- votes_processed %>% 
  group_by(year) %>%
  summarise(total = n(), percent_yes = mean(vote == 1))
by_year
```



```{r}
# Summarize by country: by_country
by_country <- votes_processed %>%
  group_by(country) %>%
  summarise(total = n(),
            percent_yes = mean(vote == 1))

by_country
```



```{r}

# Sort in ascending order of percent_yes
by_country %>% 
arrange(percent_yes)

# Now sort in descending order
by_country %>% 
arrange(desc(percent_yes))

# Filter out countries with fewer than 100 votes
by_country %>%
  arrange(percent_yes) %>% 
  filter(total >= 100)


```


```{r}
# Change to scatter plot and add smoothing curve
ggplot(by_year, aes(year, percent_yes)) +
  geom_point() + geom_smooth()
  
```

%in% will help us filter by country


```{r}
# Group by year and country: by_year_country
by_year_country <- votes_processed %>%
  group_by(year, country) %>%
  summarise(total = n(),
            percent_yes = mean(vote == 1))

# Print by_year_country
by_year_country

# Create a filtered version: UK_by_year
UK_by_year <- by_year_country %>% 
filter(country %in% "United Kingdom")

UK_by_year

# Line plot of percent_yes over time for UK only
ggplot(UK_by_year, aes(x = year, y = percent_yes)) + 
geom_line()

```



```{r}
countries <- c("United States", "United Kingdom",
               "France", "India")

by_year_country %>% 
  filter(country %in% countries) %>% 
  ggplot(aes(x = year, y = percent_yes, color = country)) + 
  geom_line()
```

With more countries it becomes to hard to understand. A better option is to create a facet for every country.

```{r}
# Vector of six countries to examine
countries <- c("United States", "United Kingdom",
               "France", "Japan", "Brazil", "India")

# Filtered by_year_country: filtered_6_countries
filtered_6_countries <- by_year_country %>% 
filter(country %in% countries) 

# Line plot of % yes over time faceted by country
ggplot(filtered_6_countries, aes(year, percent_yes)) +
  geom_line() +
  facet_wrap(~ country, scale = "free_y")
```



```{r}
# Add three more countries to this list
countries <- c("United States", "Colombia", "Dominican Republic", "Mexico")

# Filtered by_year_country: filtered_countries
filtered_countries <- by_year_country %>%
  filter(country %in% countries)

# Line plot of % yes over time faceted by country
ggplot(filtered_countries, aes(year, percent_yes)) +
  geom_line() +
  facet_wrap(~ country, scales = "free_y")

```


Now lets fit a linear model on a few countries and visualize them.

```{r}
US_by_year <- by_year_country %>% 
  filter(country == "United States")

US_fit <- lm(percent_yes ~ year, data = US_by_year)

summary(US_fit)
```


Using broom package to tidy up the summary tables into tidy dataframes.

```{r}
library(broom)

tidy(model)

# Fit model for the United Kingdom
UK_by_year <- by_year_country %>%
  filter(country == "United Kingdom")
UK_fit <- lm(percent_yes ~ year, UK_by_year)

UK_tidied <- tidy(UK_fit)
US_tidied <- tidy(US_fit)

# Combine the two tidied models
bind_rows(UK_tidied, US_tidied)

by_year_country %>% 
  ggplot(aes(x = year, y = percent_yes)) + 
  geom_line() + geom_smooth(method = "lm", se = FALSE) 


```


```{r}
#Nesting
nested <- by_year_country %>% 
  nest(-country)

nested$data[[1]]

unnest(nested)


library(broom)
library(purrr)
library(tidyr)

# Add another mutate that applies tidy() to each model
by_year_country %>%
  nest(-country) %>%
  mutate(model = map(data, ~ lm(percent_yes ~ year, data = .))) %>% 
  mutate(tidied = map(model, tidy))

# Add one more step that unnests the tidied column
country_coefficients <- by_year_country %>%
  nest(-country) %>%
  mutate(model = map(data, ~ lm(percent_yes ~ year, data = .)),
         tidied = map(model, tidy)) %>%
  unnest(tidied)

# Print the resulting country_coefficients variable
country_coefficients

# Filter for only the slope terms
country_coefficients %>% 
filter(term == "year")
```


```{r}

# Filter for only the slope terms
slope_terms <- country_coefficients %>%
  filter(term == "year")

# Add p.adjusted column, then filter
slope_terms %>%
  mutate(p.adjusted = p.adjust(p.value)) %>%
  filter(p.adjusted < .05)
# Filter by adjusted p-values
filtered_countries <- country_coefficients %>%
  filter(term == "year") %>%
  mutate(p.adjusted = p.adjust(p.value)) %>%
  filter(p.adjusted < .05)

# Sort for the countries increasing most quickly
filtered_countries %>% 
arrange(desc(estimate))

# Sort for the countries decreasing most quickly
filtered_countries %>% 
arrange(estimate)

votes_joined <- votes_processed %>% 
  inner_join(descriptions, by = c("rcid", "session"))

votes_joined %>% 
  filter(hr == 1, country == 'United States')

dominican_republic <- votes_joined %>% 
  filter(country == "Dominican Republic")
```

me: Palestinian conflict
nu: Nuclear weapons and nuclear material
di: Arms control and disarmament
hr: Human rights
co: Colonialism
ec: Economic development

```{r}
# Filter for votes related to colonialism
colonialism <- votes_joined %>% 
filter(co == 1)


# Filter, then summarize by year: US_co_by_year
US_co_by_year <- votes_joined %>%
  filter(country == "United States", co == 1) %>%
  group_by(year) %>%
  summarize(percent_yes = mean(vote == 1))

# Graph the % of "yes" votes over time
ggplot(US_co_by_year, aes(year, percent_yes)) +
  geom_line()
```


tidying up the dataframe


```{r}
votes_gathered <- votes_joined %>%
  gather(topic, has_topic, me:ec) %>% 
  filter(has_topic == 1)

votes_tidied <- votes_gathered %>%
  mutate(topic = recode(topic,
                        me = "Palestinian conflict",
                        nu = "Nuclear weapons and nuclear material",
                        di = "Arms control and disarmament",
                        hr = "Human rights",
                        co = "Colonialism",
                        ec = "Economic development"))

# Print votes_tidied
votes_tidied

# Summarize the percentage "yes" per country-year-topic
by_country_year_topic <- votes_tidied %>%
  group_by(country, year, topic) %>%
  summarize(total = n(), percent_yes = mean(vote == 1)) %>%
  ungroup()

# Print by_country_year_topic
by_country_year_topic


# Filter by_country_year_topic for just the US
US_by_country_topic <- by_country_year_topic %>% 
filter(country == "United States")

DR_by_country_year_topic <- by_country_year_topic %>% 
  filter(country == "Dominican Republic")
Col_by_country_year_topic <- by_country_year_topic %>% 
  filter(country == "Colombia")


# Plot % yes over time for the US, faceting by topic
ggplot(US_by_country_topic, aes(year, percent_yes)) + 
geom_line() + facet_wrap(~ topic)

ggplot(DR_by_country_year_topic, aes(year, percent_yes)) + 
  geom_line() + 
  facet_wrap( ~ topic) + 
  labs(x = "Time", y = "% of yes votes", title = "% of the DR voting yes to various topics over time", subtitle = "overall, there is an increase")
```


```{r}
ggplot(Col_by_country_year_topic, aes(year, percent_yes)) + 
  geom_line() + 
  facet_wrap( ~ topic) + 
  labs(x = "Time", y = "% of yes votes", title = "% of Colombia voting yes to various topics over time", subtitle = "overall, there is an increase")
```


fitting a model with year and topic

```{r}
# Load purrr, tidyr, and broom
library(purrr)
library(tidyr)
library(broom)

# Print by_country_year_topic
by_country_year_topic

# Fit model on the by_country_year_topic dataset
country_topic_coefficients <- by_country_year_topic %>% 
  nest(-country, -topic) %>% 
  mutate(models = map(data, ~ lm(percent_yes ~ year, data = .))) %>% 
  mutate(tidied = map(models, tidy)) %>% 
  unnest(tidied)

# Print country_topic_coefficients
country_topic_coefficients


# Create country_topic_filtered
#filter by significant coefficients

country_topic_filtered <- country_topic_coefficients %>% 
  filter(term == "year") %>% 
  mutate(p.adjusted = p.adjust(p.value)) %>% 
  filter(p.adjusted < 0.05)

country_topic_filtered %>% 
  arrange(desc(estimate))

country_topic_filtered %>% #with significant coefficients
  arrange(estimate)


# Create vanuatu_by_country_year_topic
vanuatu_by_country_year_topic <- by_country_year_topic %>%
  filter(country == "Vanuatu")

vanuatu_by_country_year_topic
# Plot of percentage "yes" over time, faceted by topic
ggplot(vanuatu_by_country_year_topic, aes(year, percent_yes)) +
  geom_line() +
  facet_wrap(~ topic)
```


```{r}



```

